Customer Segmentation

### 2024-03-17 R/eda-code/eda_superstore.qmd
Project Part 3

#| message: false
#| warning: false
#load needed packages. 
library(here) #for data loading/saving
library(dplyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(ggplot2) # for visualization
library(tibble) # For rownames_to_column
library(gt) # for table
library(gridExtra)
library(webshot2) #Capture as an Image the gttable before converting it to pdf
```

@@

```{r}
#Path to data using here() package
data_location <- here::here("data","processed-data","processed_superstore.rds")
#load data
mydata <- readRDS(data_location)
```

@@

It is important to note that this data accounts sales of a particular item at a time as an observation leading different items sold in the same Invoice to appear as separate observations. As such most of the Invoice numbers appear several times in different observations of the data. A meaningful analysis of customer spending pattern require the Invoice Sales Amount, total purchase by a Customer during a period, frequency of purchase by a customer during a period or a combination of those. A table is created which provides the descriptive statistics of sales by Invoice value, Customer purchase and frequency of customer purchase. This table reveals that there were 1479 transactions to 414 different customers during 2010-12-01 and	2011-12-09.

@@

# Aggregate total sales by InvoiceNo for calculating mean, standard deviation, minimum and maximum
Sales_by_Invoice <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  ungroup() %>% # Ensure we ungroup before further summarisation
  summarise(
    Count = n(),
    Particulars = "Sales by Invoice",
    Mean = round(mean(TotalSales, na.rm = TRUE), 2),
    SD = round(sd(TotalSales, na.rm = TRUE), 2),
    Min = round(min(TotalSales, na.rm = TRUE), 2),
    Max = round(max(TotalSales, na.rm = TRUE), 2)
  )

# Aggregate total sales by CustomerID to calculate mean, standard deviation, minimum and maximum
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(
    Count = n(),
    Particulars = "Cumulative Sales to Individual Customers",
    Mean = round(mean(TotalSales, na.rm = TRUE), 2),
    SD = round(sd(TotalSales, na.rm = TRUE), 2),
    Min = round(min(TotalSales, na.rm = TRUE), 2),
    Max = round(max(TotalSales, na.rm = TRUE), 2)
  )

@@

```{r}
# Aggregate sales by CustomerID
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(Sales, na.rm = TRUE)) %>%
  arrange(desc(TotalSales))

@@

# Histogram of Sales by Invoice values
## First calculate total invoice value per invoice
Sales_by_Invoice <- mydata %>%
  group_by(InvoiceNo) %>%
  summarise(TotalInvoiceValue = sum(Sales, na.rm = TRUE))

##Create histogram
p2 <- ggplot(Sales_by_Invoice, aes(x = TotalInvoiceValue)) +
  geom_density(fill = "lightgreen", alpha = 0.5) +
  labs(title = "Density Plot of  Invoice Sales",
       x = "Invoice Value", 
       y = "Density") +
  theme_minimal()

@@

A histogram is also plotted for price range of the products.Most of the products fall below the price of 12.50 sterling pounds.

```{r}
# Create histogram of Unit Price

p4 <- ggplot(mydata, aes(x = UnitPrice)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "darkblue") + # Adjust binwidth as needed
  labs(title = "Histogram of Unit Price",
       x = "Unit Price", 
       y = "Frequency") +
  theme_minimal()

p4

figure_file = here("results","figures","Price_range.png")
ggsave(filename = figure_file, plot=p4)

```
Next, total purchase by a customer was plotted against purchase frequency.A few outliers were noted.

```{r}
#| message: false
#| warning: false
# Calculate purchase frequency and total invoice value per customer
customer_stats <- mydata %>%
  group_by(CustomerID) %>%
  summarise(PurchaseFrequency = n_distinct(InvoiceNo),
            TotalInvoiceValue = sum(Sales, na.rm = TRUE)) %>%
  ungroup()  # Ensure the data is ungrouped for further analysis

# Scatter plot of Total Invoice Value by Purchase Frequency
p5 <- ggplot(customer_stats, aes(x = PurchaseFrequency, y = TotalInvoiceValue)) +
  geom_point(alpha = 0.5) +  # Adjust alpha for point transparency, if needed
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without standard error
  labs(title = "Total Invoice Value by Purchase Frequency",
       x = "Purchase Frequency",
       y = "Total Invoice Value") +
  theme_minimal()
plot(p5)

figure_file = here("results","figures","Sales_trend.png")
ggsave(filename = figure_file, plot=p5)

@@

```{r}
p8 <- mydata %>% 
  group_by(Country) %>% 
  summarise(TotalSales = sum(Sales)) %>%
  ggplot(aes(x = reorder(Country, TotalSales), y = TotalSales, fill = Country)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  theme(legend.position = "none")

plot(p8)

figure_file = here("results","figures","Countrywise_Sales.png")
ggsave(filename = figure_file, plot=p8)
```

>>>>

#| message: false
#| warning: false
#load needed packages. 
library(here) #for data loading/saving
library(dplyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(ggplot2) # for visualization
library(tibble) # For rownames_to_column
library(gt) # for table
library(gridExtra)
library(webshot2) #Capture as an Image the gttable before converting it to pdf
library(zoo) #To plot seasonality of sales
```

@@

```{r}
#Path to data using here() package
#data_location <- here::here("data","processed-data","processed_superstore.rds")
data_location <- here::here("data","processed-data","processed_superstore_RFM.rds")
#load data
mydata <- readRDS(data_location)
```

@@

A meaningful analysis of customer spending behavior require the Invoice Sales Amount, total purchase by a Customer during a period, frequency of purchase by a customer during a period or a combination of those. A table is created which provides the descriptive statistics of sales by Invoice value, Customer purchase and frequency of customer purchase. This table reveals that there were 1402 transactions to 399 different customers during 2010-12-01 and	2011-12-09.

@@

# Aggregate total sales by InvoiceNo for calculating mean, standard deviation, minimum and maximum
Sales_by_Invoice <- mydata %>%
    summarise(
    Count = n(),
    Particulars = "Invoice Value",
    Mean = round(mean(InvoiceValue, na.rm = TRUE), 2),
    SD = round(sd(InvoiceValue, na.rm = TRUE), 2),
    Min = round(min(InvoiceValue, na.rm = TRUE), 2),
    Max = round(max(InvoiceValue, na.rm = TRUE), 2)
  )

# Aggregate total sales by CustomerID to calculate mean, standard deviation, minimum and maximum
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(InvoiceValue, na.rm = TRUE)) %>%
  ungroup() %>%
  summarise(
    Count = n(),
    Particulars = "Cumulative Sales to Individual Customers",
    Mean = round(mean(TotalSales, na.rm = TRUE), 2),
    SD = round(sd(TotalSales, na.rm = TRUE), 2),
    Min = round(min(TotalSales, na.rm = TRUE), 2),
    Max = round(max(TotalSales, na.rm = TRUE), 2)
  )

@@

```{r}
# Aggregate sales by CustomerID
TotalSales_customer <- mydata %>%
  group_by(CustomerID) %>%
  summarise(TotalSales = sum(InvoiceValue, na.rm = TRUE)) %>%
  arrange(desc(TotalSales))

@@

# Histogram of Sales by Invoice values

p2 <- ggplot(mydata, aes(x = InvoiceValue)) +
  geom_density(fill = "lightgreen", alpha = 0.5) +
  labs(title = "Density Plot of  Invoice Sales",
       x = "Invoice Value", 
       y = "Density") +
  theme_minimal()

@@

Next, total purchase by a customer was plotted against purchase frequency.A few outliers were noted.

```{r}
#| message: false
#| warning: false
# Calculate purchase frequency and total invoice value per customer
customer_stats <- mydata %>%
  group_by(CustomerID) %>%
  summarise(PurchaseFrequency = n_distinct(InvoiceNo),
            TotalInvoiceValue = sum(InvoiceValue, na.rm = TRUE)) %>%
  ungroup()  # Ensure the data is ungrouped for further analysis

# Scatter plot of Total Invoice Value by Purchase Frequency
p4 <- ggplot(customer_stats, aes(x = PurchaseFrequency, y = TotalInvoiceValue)) +
  geom_point(alpha = 0.5) +  # Adjust alpha for point transparency, if needed
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without standard error
  labs(title = "Total Invoice Value by Purchase Frequency",
       x = "Purchase Frequency",
       y = "Total Invoice Value") +
  theme_minimal()
plot(p4)

figure_file = here("results","figures","Sales_trend.png")
ggsave(filename = figure_file, plot=p4)

@@

```{r}
p5 <- mydata %>% 
  group_by(Country) %>% 
  summarise(TotalSales = sum(InvoiceValue)) %>%
  ggplot(aes(x = reorder(Country, TotalSales), y = TotalSales, fill = Country)) + 
  geom_bar(stat = "identity") + 
  coord_flip() + 
  theme(legend.position = "none")

plot(p5)

figure_file = here("results","figures","Countrywise_Sales.png")
ggsave(filename = figure_file, plot=p5)
```
As the store sells gift goods, an inspection is required to check the seasonality of sales. First, monthly total sales is calculated aggregating Invoice Value by months of the Invoice Date.The aggregated sales is then plotted to visualize seasonal trend which depicted a spike in the sales during the last few months of the year.

```{r}
#Aggregating Sales by months
mydata_aggregated <- mydata %>%
  group_by(InvoiceDate = as.yearmon(InvoiceDate)) %>%
  summarize(TotalValue = sum(InvoiceValue))

# Formatting Date for plotting purposes
mydata_aggregated$InvoiceDate <- as.Date(as.yearmon(mydata_aggregated$InvoiceDate))

p6<-ggplot(mydata_aggregated, aes(x = InvoiceDate, y = TotalValue)) +
  geom_line() + 
  scale_x_date(date_labels = "%b", date_breaks = "1 month") + # Adjust formatting as needed
  labs(title = "Monthly Sales", x = "Month", y = "Sales Value") +
  theme_minimal()
p6

figure_file = here("results","figures","Monthly_Sales.png")
ggsave(filename = figure_file, plot=p6)

### 2024-03-17 R/processing-code/processing_superstore-v1.qmd
Project Part 3

@@

```{r}
processed_superstore <- clean_data_step4

str(processed_superstore)
```

@@

>>>>

The wholesalers are expected to purchase in large quantity. Additional inspection is carried out to identify wholesalers by quantity purchased and individual Purchase Amount by plotting a scatter plot.

```{r}
# Calculate purchase frequency and total invoice value per customer
customer_stats5 <- clean_data_step4 %>%
  group_by(InvoiceNo) %>%
  summarise(InvoiceQty = sum(Quantity, na.rm = TRUE),
            InvoiceValue = sum(Sales, na.rm = TRUE)) %>%
  ungroup()  # Ensure the data is ungrouped for further analysis

# Scatter plot of Total Invoice Value by Purchase Frequency
p3 <- ggplot(customer_stats5, aes(x = InvoiceValue, y = InvoiceQty)) +
  geom_point(alpha = 0.5) +  # Adjust alpha for point transparency, if needed
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without standard error
  labs(title = "Total Invoice Value by Purchase Frequency",
       x = "Invoice Amount",
       y = "Quantity Purchased") +
  scale_y_continuous(labels = label_comma()) +  # This line changes the y-axis labels

    theme_minimal()
plot(p3)

```
The scatter plot reveals that most of the customers made purchases less than 1000 items and most of the invoice values were less than 2,500. Assuming that the sparsely distributed points represent purchases by wholesalers, observations with Quantity above 1000 and individual Invoice value above 2,500 are removed from the data.

This involves first sub-setting customer_stats5 such that it includes only those observations where InvoiceQty <= 1000 and InvoiceValue <=2500 and finally, merging this sub-set with the original dataset (clean_data_step4).

```{r}
# Filtering the customer_stats5 to meet the conditions
filtered_customer_stats5 <- customer_stats5 %>%
  filter(InvoiceQty <= 1000, InvoiceValue <= 2500)

# Excluding purchase quantities above 1000 and individual invoice values above 2500 from the data
clean_data_step5 <- clean_data_step4 %>%
  semi_join(filtered_customer_stats5, by='InvoiceNo')

skimr::skim(clean_data_step5)
```
This step excluded 4485 observations from the data for a final total observation of 28117. A similar scatter plot is plotted for the new data.

```{r}
# Calculate purchase quantity and amount per invoice
customer_stats6 <- clean_data_step5 %>%
  group_by(InvoiceNo) %>%
  summarise(InvoiceQty = sum(Quantity, na.rm = TRUE),
            InvoiceValue = sum(Sales, na.rm = TRUE)) %>%
  ungroup()  # Ensure the data is ungrouped for further analysis

# Scatter plot of Purchase Quantity by Invoice Value
p4 <- ggplot(customer_stats6, aes(x = InvoiceValue, y = InvoiceQty)) +
  geom_point(alpha = 0.5) +  # Adjust alpha for point transparency, if needed
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear regression line without standard error
  labs(title = "Purchase Quantity by Invoice Value",
       x = "Invoice Amount",
       y = "Quantity Purchased") +
  scale_y_continuous(labels = label_comma()) +  # This line changes the y-axis labels

    theme_minimal()
plot(p4)

@@

```{r}
processed_superstore <- clean_data_step5

str(processed_superstore)
```
It is important to note that this data accounts sales of a particular item at a time as an observation leading different items sold in the same Invoice to appear as separate observations. As such most of the Invoice numbers appear several times in different observations of the data. A meaningful analysis of customer spending pattern can be done by total Invoice Sales. So another data is created where each Invoice is a unique observation and can be used for groupoing customers based on Recency, Frequency and Monetary value.

```{r}
processed_superstore_RFM <- processed_superstore %>%
  group_by(InvoiceNo) %>%
  summarise(InvoiceDate = first(InvoiceDate),
            CustomerID = first(CustomerID),
            InvoiceValue = sum(Sales, na.rm = TRUE),
            Country = first(Country)) %>%
select(InvoiceDate, InvoiceNo, CustomerID, InvoiceValue, Country)

str(processed_superstore_RFM)
```

@@

```{r}
save_data_location <- here::here("data","processed-data","processed_superstore_RFM.rds")
saveRDS(processed_superstore_RFM, file = save_data_location)

### 2024-03-17 products/manuscript/Manuscript.qmd
Project Part 3

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

@@

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

Finite mixture models are used to analyze data presumably arising from several separate, yet unobserved, groups. Among these models, Gaussian mixture models and mixture of t-distributions are particularly prevalent. However, such models which are based on symmetric distribution often do not accurately represent data with asymmetric distribution. Skewed data can be more effectively managed using asymmetric distributions, thereby extending their applicability across a broader spectrum of model-based clustering scenarios [1]. If the observed data is positive, distribution is right-skewed and assumed to come from a mixture distribution, employing mixture of gamma density is a rational choice [2].

This study intends to identify distinct customer groups within sales data of luxury goods, with each group comprising customers who exhibit similar characteristics in terms of purchase frequency and amounts. Modeling customer spending as a mixture of gamma distributions helps to identify different spending behaviors. The objective is to develop an Expectation Maximization (EM) algorithm which segregates groups based on their spending patterns, considering the individual purchases transactions made by each customer.

@@

This study employs simulated data to assess the effectiveness of the EM algorithm. Following successful testing, the refined algorithm will be applied to actual sales data pertaining to luxury goods.

@@

This study aims to answer the question, "How effective is the Expectation Maximization (EM) algorithm in identifying distinct customer groups within luxury goods sales data when accounting for individual spending by each customer, as demonstrated by initial validation using simulated data?"

[1] Lee, S. X., & McLachlan, G. J. (2013). Model-based clustering and classification with non-normal mixture distributions. Statistical Methods & Applications, 22(4), 427-454.

[2] Young, D. S., Chen, X., Hewage, D. C., & Nilo-Poyanco, R. (2019). Finite mixture-of-gamma distributions: estimation, inference, and model-based clustering. Advances in Data Analysis and Classification, 13, 1053-1082.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

@@

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
#knitr::include_graphics(here("products","assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

@@

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

@@

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","height-weight-stratified.png"))
```

@@

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables-files","resulttable2.rds"))
knitr::kable(resulttable2)
```

>>>>

---
title: "Customer Segmentation"
subtitle: ""
author: Malika Dhakhwa
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

@@

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis

This project aims to explore customer segmentation through the supervised and un-supervised learning perspectives. In the initial phase, the analysis applies and investigates the effectiveness of logistic regression to segregate customers based on the general patterns of customer behavior. As the study advances, it integrates more advanced tools, such as Expectation Maximization algorithms.

It is crucial that organizations understand their customer’s purchasing patterns to design and execute effective and customized  engagement strategies. This helps businesses enhance customer service, consequently increase sales. Based on characteristics like spending and purchase frequency, businesses can segment customers and tailor their marketing and sales efforts to meet the unique needs of each group. This personalized approach can enhance customer satisfaction and also improve multiple end results like customer loyalty, operational efficiency, and profitability.

The original data does not have the customers categorized into a group. To use logistic regression, a group variable is created based on the purchase amount and frequency.

@@

This study utilizes the 'superstoredata'dataset, focusing on international sales from an online retail store.

Initially this data was downloaded from Kaggle for a preliminary assessment of it's usability for this project. However, when revisiting the site for further details, it was no longer available on Kaggle.com. A subsequent Google search led to the discovery of the dataset in the [UC Irvine Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/Online+Retail). Per the repository, this is a transnational data set spanning from December 1, 2010, to December 9, 2011 from a UK-based, non-store online retail operation. The company specializes in unique all-occasion gifts to a clientele which also includes wholesalers.

This data has nine variables as detailed below:

(1)	InvoiceNo: It is a categorical variable with a 6-digit integral number uniquely assigned to each transaction. If this code starts with the letter ‘c’, it indicates a cancellation. 
(2)	StockCode: It is a categorical variable with a 5-digit integral number uniquely assigned to each distinct product.
(3)	Description: This represents Product name and is a categorical variable.
(4)	Quantity: It represents quantities of each product (item) per transaction and is an integer.
(5)	InvoiceDate: It represents the day and time each transaction was generated.
(6)	UnitPrice: It represents product price per unit in sterling.
(7)	CustomerID: It is a categorical variable with a 5-digit integral number uniquely assigned to each customer.
(8)	Country: It represents the name of the country where each customer resides.
(9) Sales: It represents product price multiplied by UnitPrice.

@@

The objective is to understand customer segmentation from logistic regression and advanced analytical tools such as Mixture Models.

@@

## Schematic of workflow
_Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). is an example of some - completely random/unrelated - schematic that was generated with Biorender. We store those figures in the assets folder.

@@

The primary dataset for this project is the 'superstoredata' dataset, detailing transactions of an online retailer based in United Kingdom. A key feature of this data is that it counts sales of a particular item from an Invoice as an observation leading to different items sold in the same Invoice to appear as separate observations. Most of the Invoice numbers appear several times in different data observations.

### Initial Data Cleaning

As the project focuses on international customers, initial data cleaning involved eliminating UK transactions downsizing the data size to 46431 observations across nine variables.

The preliminary inspection identified several issues needing attention such as missing values in 'CustomerID', negative 'signs in 'Quantity' and 'Sales' values, and incorrectly formatted dates. Variables, 'InvoiceNo', 'StockCode', 'Description' and 'Country', originally coded as characters, were converted to factor types to facilitate better data manipulation.

1480 observations lacked 'CustomerID' and were removed, as customer behavior cannot be tracked without the unique identifier and 'CustomerID' was converted to factor type from numeric type. 'Country', 'Description' and 'StockCode' were converted to factors from character variables and transaction date to the date format.

There were 1372 observations with negative values in quantity and sales due to order cancellation which were excluded from the data.

An examination of 'UnitPrice' distribution revealed that all the high value goods with a Unitprice above 25 were distributed between 75 and 100 percentile. A closer inspection at this inter-quartile range disclosed that there were 19 product description categories with the Unitprice above 25. Some of those categories had descriptions such as CARRIAGE, POSTAGE, Manual, which appeared to be related to logistics. 1216 observations with such descriptions were dropped from the data since those were less likely to be directly related to customer purchase.

### Refining the Dataset
The store also sells merchandises in wholesale quantities. A wholesaler potentially had larger purchases and in higher frequencies. In absence of clear information on wholesalers, a scatter plot of Cumulative Invoice Value of customers by Purchase Frequency was plotted to identify potential wholesalers. The scatter plot revealed that most of the customers made purchases less than a frequency of 50 and that most cumulative purchase values were less than 50,000. Assuming that the outliers represented purchases by wholesalers, observations with purchase frequencies above 50 and cumulative Invoice values above 50,000 were dropped from the data. This step excluded 9745 observations from the data. The cleaned data has 32602 observations with 9 variables. Next, the quantities purchased in each invoice were plotted against the individual invoice value. The scatter plot revealed that the observations were concentrated within the purchase quantity below 1000 items and invoice amounts of 2,500. Assuming that the sparsely distributed points represent purchases by wholesalers, observations with Quantity above 1000 and individual Invoice value above 2,500 were further removed from the data.This step excluded 4485 observations for a final total observation of 28117.

The study of customer spending pattern requires the knowledge of each individual transaction amount. The subsequent cleaning process reorganized the dataset by 'InvoiceNo', consolidating sales data under each invoice. This final data is saved as processed_superstore_RFM in the data > processed-data.

### RFM Classification
In absence of a labeled categorical variable for classification, the study segmented customers into two distinct groups using a recency, frequency and monetary (RFM) model. RFM model is a commonly used behavior-based model in analysing customer behavior [@yeh2009knowledge]. Recency is measured in days or months and represent the interval between the most recent transaction time and the time of analysis, and a lower number interval is preferred. Frequency is the number of purchases made in a certain period, and monetary is the total amount the customer spent during that time period.[@wei2010review]

In the RFM classification process, the Recency, Frequency and Monetary values were calculated for each customer, ensuring that each customer was represented only once in the modified dataset.Frequency in this data is the total no. of Invoices issued to the customer, and Monetary is the total purchases made by the customer during the stipulated period. Recency was calculated from the next day of the transaction period. Each customer was assigned a value ranging from 1 to 4 for the Recency, Frequency and Monetary attributes, reflecting their quartile position within each metric. Subsequently, these metrics were combined to formulate a three-digit score for every custoemr, positioning the Monetary value in the hundreds place, Frequency in the tens place, and Recency in the units place. All the customers with a score above 344 were assigned as High-Value Customers.

Following RFM segmentation, logistic regression was applied to predict customer group membership. This method utilized the segments derived from RFM as the dependent variable, with the Recency, Frequency and Monetary attritubtes serving as independent variables. The objective is to develop a model for predictive analysis based on those segments.

@@

@tbl-resulttable1 depicts the descriptive statistics of various aspects of customers purchase behavior. There were 1402 transactions with 399 customers outside UK during the period from December 1, 2010 to	December 9, 2011. The store recorded a mean sales of GBP 436 per transaction, an average cumulative sales of GBP 1532 to individual customers and sales frequency of above 3 times per customer.

```{r out.width="100%", fig.show='hold', warning=FALSE}
#| label: tbl-resulttable1
#| tbl-cap: "Descriptive Statistics"
#| echo: FALSE
knitr::include_graphics(here("results","tables", "gt_table.png"))

```

@fig-result1 depicts the distribution of customer segmentation and suggests that approximately 25 percentage of the customers are grouped as Valued Customers.

```{r}
#| label: fig-result1
#| fig-cap: "Distribution of Sales and Purchase frequency"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Customer_Segment.png"))
```

@fig-result2 depicts two density plots for the Cumulative Sales and Invoice Values across customers and a histogram for Customer purchase frequencies. Both density plots are heavily right-skewed.T0he averages of total sales to a customer and total invoice values appear to be less than 2500 and 500 sterling pounds, respectively, supporting the mean values of 1532 and 436 in the descriptive table. The average frequency appears to be less than 5 times, which is also observed in the descriptive statistics table.

```{r}
#| label: fig-result2
#| fig-cap: "Distribution of Sales and Purchase frequency"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Sales_distribution.png"))
```
@fig-result3 depicts a density plot of number of days since the last transaction made by a customer. The distribution is right-skewed. This plot indicates that most customers made a purchase transaction within the last 30 days of the reference date.It also reveals that few customers had not make a transaction for a long time.

```{r}
#| label: fig-result3
#| fig-cap: "Distribution of Recency of Purchase"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Recency.png"))
```

@fig-result4 depicts a trend of monthly international sales. A spike is noticed in November, presumably due to Christmas in the following month, which is also supported by distribution of the Recency plot. 
```{r}
#| label: fig-result4
#| fig-cap: "Monthly Sales trend"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Monthly_Sales.png"))
```
@fig-result5 depicts country-wise trend of international sales by sales amount. Top customers were in Germany, France, Spain and Belgium.  
```{r}
#| label: fig-result5
#| fig-cap: "Countrywise Sales trend"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "Countrywise_Sales.png"))
```

@@

Penalized logistic regression of customer segment was performed on Recency, Frequency and Monetary variables using 'tidymodels' package and 4-fold corss-validation techniques.

@fig-result6 depicts the area under ROC curve by the range of penalty values. It provides visualization of validation set metrics. The curve has extreme high values, all above 0.9995. A sharp drop is noticed before the penalty value of 0.0010, suggesting the sensitivity of the performance of the model at this penalty value. This is followed by a sharp rise after the penalty value of 0.01.  These features of model performance are indicative of model overfitting to the training data and the data being easily separable.

```{r}
#| label: fig-result6
#| fig-cap: "Visualization of validation set metrics"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "ROC(AUC)_Penalty.png"))
```
@fig-result7 depicts validation set ROC curve plotting true positive rate (sensitivity) against the false positive rate (1-specificity) at various threshold settings. The curve moves along 1-specificity and rises only from the end.This suggests that the model has no  ability to distinguish between the two customer categories. This suggests that the logistic regression model is not appropriate to the dataset. It pointed out the need to revisit the choice of model.

```{r}
#| label: fig-result7
#| fig-cap: "Visualization of Validation set ROC curve"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "ROC(AUC)_Validation.png"))
```

@@

#Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
#resulttable2 = readRDS(here("results","tables","summarytable.rds"))
#knitr::kable(resulttable2)
```
